---
title: "Creating graphs in R with `ggplot2`"
author: "Luca"
date: "03/12/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

# Load packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

# Load files
```{r}
ca_np <- read_csv(here("data", "ca_np.csv"))
```

# Look at the data
```{r}
names(ca_np)
head(ca_np)
summary(ca_np)
```

```{r}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

```{r}
names(ci_np)
head(ci_np)
summary(ci_np)
```

To create a bare-bones ggplot graph, we need to tell R three basic things:

1. We’re using ggplot2::ggplot()
2. Data we’re using & variables we’re plotting (i.e., what is x and/or y?)
3. What type of graph we’re making (the type of geom)

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

We’re going to be doing a lot of plot variations with those same variables. Let’s store the first line as object gg_base so that we don’t need to retype it each time:

```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
```

Change to a scatterplot
```{r}
gg_base + geom_point()
```

Change to a column graph
```{r}
gg_base + geom_col()
```

Change to an area plot
```{r}
gg_base + geom_area()
```

First, we’ll customize some aesthetics (e.g. colors, styles, axis labels, etc.) of our graphs based on non-variable values.

We can change the aesthetics of elements in a ggplot graph by adding arguments within the layer where that element is created.

Some common arguments we’ll use first are:

`color` = or colour =: update point or line colors
`fill` =: update fill color for objects with areas
`linetype` =: update the line type (dashed, long dash, etc.)
`pch` =: update the point style
`size` =: update the element size (e.g. of points or line thickness)
`alpha` =: update element opacity (1 = opaque, 0 = transparent)

Building on our first line graph, let’s update the line color to “purple” and make the line type “dashed”:

```{r}
gg_base +
  geom_line(
    color = "purple",
    linetype = "dashed"
  )
```

Now let’s update the point, style and size of points on our previous scatterplot graph using `color =`, `size =`, and `pch =` (see ?pch for the different point styles, which can be further customized).

```{r}
gg_base + 
  geom_point(color = "purple",
             pch = 17,
             size = 4,
             alpha = 0.5)
```

Sometimes, however, we do want the aesthetics of a graph to depend on a variable. To do that, we’ll map variables onto graph aesthetics, meaning we’ll change how an element on the graph looks based on a variable characteristic (usually, character or value).

When we want to customize a graph element based on a variable’s characteristic or value, add the argument within `aes()` in the appropriate `geom_*()` layer

Example: Create a ggplot scatterplot graph where the size and color of the points change based on the number of visitors, and make all points the same level of opacity (alpha = 0.5). Notice the aes() around the size = and color = arguments.

```{r}
gg_base + 
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  )
```

notice that the two arguments that do depend on variables are within `aes()`, but since `alpha = 0.5` doesn’t depend on a variable then it is outside the `aes()` but still within the `geom_point()` layer.


Using themes
```{r}
gg_base +
  geom_point(
    aes(size = visitors,
        color = visitors),
    alpha = 0.5
  ) +
  theme_minimal()
```


Use `labs()` to update axis labels, and add a title and/or subtitle to your ggplot graph.
```{r}
gg_base +
  geom_line(linetype = "dotted") +
  theme_minimal() +
  labs(
    x = "Year",
    y = "Annual park visitors",
    title = "Channel Islands NP Visitation",
    subtitle = "(1963 - 2016)"
  )
```


As long as the geoms are compatible, we can layer them on top of one another to further customize a graph.

For example, adding points to a line graph:

```{r}
gg_base +
  geom_line(color = "purple") +
  geom_point(color = "orange",
             aes(size = year),
             alpha = 0.5)
```

Often we’ll want to visualize multiple series. For example, from the ca_np object we have stored, we might want to plot visitation for all California National Parks.

To do that, we need to add an aesthetic that lets `ggplot` know how things are going to be grouped. A demonstration of why that’s important - what happens if we don’t let ggplot know how to group things?

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors)) +
  geom_line()
```

Well that’s definitely a mess, and it’s because ggplot has no idea that these should be different series based on the different parks that appear in the ‘park_name’ column.

We can make sure R does know by adding an *explicit grouping argument `(group =)`, or by updating an aesthetic based on park_name:*

```{r}
ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name)) +
  geom_line()
```

*Note:* You could also add an aesthetic (`color = park_nam`e) in the `geom_line()` layer to create groupings, instead of in the topmost `ggplot()` layer.

Let’s store that topmost line so that we can use it more quickly later on in the lesson:

```{r}
gg_np <- ggplot(data = ca_np, aes(x = year, y = visitors, group = park_name))
```

Faceting graphs

When we facet graphs, we split them up into multiple plotting panels, where each panel contains a subset of the data. In our case, we’ll split the graph above into different panels, each containing visitation data for a single park.

Also notice that *any general theme changes made will be applied to all of the graphs*.

```{r}
gg_np +
  geom_line(show.legend = FALSE) +
  theme_light() + 
  labs(x = "year", y = "annual visitors") +
  facet_wrap(~ park_name)
```

Exporting a ggplot graph with `ggsave()`

If we want our graph to appear in a knitted html, then we don’t need to do anything else. But often we’ll need a saved image file, of specific size and resolution, to share or for publication.

`ggsave()` will export the most recently run ggplot graph by default (`plot = last_plot()`), unless you give it the name of a different saved ggplot object. Some common arguments for `ggsave()`:

* `width =`: set exported image width (default inches)
* `height =`: set exported image height (default height)
* `dpi =`: set dpi (dots per inch)
So to export the faceted graph above at 180 dpi, width a width of 8" and a height of 7", we can use:

```{r}
ggsave(here("figures", "np_graph.jpg"), dpi = 180, width = 8, height = 7)
```









