---
title: "Creating graphs in R with `ggplot2`"
author: "Luca"
date: "03/12/2021"
output: html_document
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Plots with ggplot2

We are going to make plots in R and it's going to be amazing.

# Load packages
```{r}
library(tidyverse)
library(readxl)
library(here)
```

# Load files
```{r}
ca_np <- read_csv(here("data", "ca_np.csv"))
```

# Look at the data
```{r}
names(ca_np)
head(ca_np)
summary(ca_np)
```

```{r}
ci_np <- read_excel(here("data", "ci_np.xlsx"))
```

```{r}
names(ci_np)
head(ci_np)
summary(ci_np)
```

To create a bare-bones ggplot graph, we need to tell R three basic things:

1. We’re using ggplot2::ggplot()
2. Data we’re using & variables we’re plotting (i.e., what is x and/or y?)
3. What type of graph we’re making (the type of geom)

```{r}
ggplot(data = ci_np, aes(x = year, y = visitors)) +
  geom_line()
```

We’re going to be doing a lot of plot variations with those same variables. Let’s store the first line as object gg_base so that we don’t need to retype it each time:

```{r}
gg_base <- ggplot(data = ci_np, aes(x = year, y = visitors))
```

Change to a scatterplot
```{r}
gg_base + geom_point()
```

Change to a column graph
```{r}
gg_base + geom_col()
```

Change to an area plot
```{r}
gg_base + geom_area()
```














